---
title: "Take Home Exercise 3"
author: "Law Man Long"
date: "11 February 2023"
date-modified: "`r Sys.Date()`"
execute: 
  echo: true
  eval: true
  results: hide
  message: false
  warning: false
format: html
editor: visual
---

# 1. Overview

The purpose of the exervise is to uncover the salient patterns of resale prices of public housing property by residential towns and estates in Singapore. I will use select appropriate analytical visualisation techniques to discover the similarity and difference of data pattern for 3-Room, 4-Room and 5-Room flats. The study period is base on 2022 resale data that is available on [data.gov.sg](https://data.gov.sg/dataset?organization=housing-and-development-board) named "resale-flat-prices-based-on-registration-date-from-jan-2017-onwards".

# 2. Visualization Tool Selection

### Boxplot

1\) Compare resale flat prices for 3-Room, 4-Room and 5-Room flat

2\) Compare resale flat prices between different town

3\) Compare resale flat prices between different flat model

### Correlation Analysis

1\) Scatterplot to show the relationship of floor sq and resale flat price

2\) Remaining lease years vs resale flat price

### Other Independent Analysis

1\) Mosaic plot to show the relationship of X & Y

# 3. Step by Step Procedures

3.1 Getting Ready

```{r}
pacman::p_load(tidyverse, plotly)
```

3.2 Import data

```{r}
flat_data <- read_csv("data/resale-flat-prices-jan2017.csv")
```

# Data Patterns
