---
title: "Take Home Exercise 4"
author: "Law Man Long"
date: "18 February 2023"
date-modified: "`r Sys.Date()`"
execute: 
  echo: true
  eval: true
  results: hide
  message: false
  warning: false
format: html
editor: visual
---

## 1. Overview

## 2. Data Preparation

### 2.1 Getting Ready

In this exercise, I will use ggplot2 and its extension, tidyverse, gganimate and its extension

Package

gganimate, CPGfunction (newggslopegraph), [gifski_renderer()](https://gganimate.com/reference/renderers.html)

```{r}
pacman::p_load(ggstatsplot, plotly, gganimate, readxl, tidyverse)
```

### 2.2 Import Data

I use the below code chunk to extract the data from the selected excel spreadsheets and range, and store it as "Import_trade" and "Export_trade" data frame.

```{r}
Import_trade <- read_excel("data/merchandisetrade.xlsx", sheet = "T1", range = "A10:AL129")
```

```{r}
Export_trade <- read_excel("data/merchandisetrade.xlsx", sheet = "T2", range = "A10:AL101")
```

### 2.3 Data Preparation

The dataframe is at a wider table format and it is not user-friendly for time-series analysis. The below code chunk transforms the data from wider to longer format.

```{r}
Import_trade2 <- pivot_longer(Import_trade, cols = 2:38, names_to = "Month year", values_to = "Trade Value" )
```

```{r}
Export_trade2 <- pivot_longer(Export_trade, cols = 2:38, names_to = "Month year", values_to = "Trade Value" )
```

```{r}
head(Import_trade2)
```

I then join the two data frame to one data frame and create one column for import trade value and one column for export trade value in dollar value.

```{r}
# join two data and create "import" and "export" column
```

There are 4 columns within the dataframe. For more accurate description, I rename the first column as "Country", remove the irrelevant text and the "total" row.

```{r}
# rename column name
# remove "thousand dollar"
# remove first row
```

For further analyzing time pattern, I further derive the time stamp to month and year

```{r}
# derive month, year from month year
```

## 3. Data Visualisation

### 3.1 Line graph

```{r}
# Line graph with overall import and export
# Trellis line graph to show the top 10 trade countries
```

### 3.2 Slope graph

```{r}
# slopegraph to shows the total trade volume changes for 2020 and 2022
```

### 3.3 Horizon graph.

```{r}

```

### 3.4 Cycle plot

```{r}
# cycle plot to shows the month and year pattern for overall countries and selected countries
```

### 3.5 Animated bubble plot

```{r}
# animated scatterplot with X = Import, Y = Export, size of bubble = overall trade
```

## 4. Data Insights

## 5. Learnings
